---
title: "Our World in Data"
author: "Daniel Alvarez"
date: "8/19/2021"
output: pdf_document
---

```{r Installations and  setup, include=FALSE}
#install.packages("devtools")
# install.packages(c("tidyverse","tidyr","dplyr","car","wooldridge",
#                    "lmtest","sandwich","ggplot","ggplot2",
#                    "GGally","zoo","lubridate","foreign","data.table"))
library(devtools)
devtools::install_github("piersyork/owidR")
library(tidyverse)
library(tidyr)
library(dplyr)
library(car)
library(wooldridge)
library(lmtest)
library(sandwich)
#library(ggplot)
library(ggplot2)
library(GGally)
library(zoo)
library(lubridate)
library(foreign)
library(data.table)

library(owidR) #Our World in Data library

tinytex::install_tinytex()
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=70, 
                                                  tidy=TRUE))
```

## Example of using Our World in Data

This package acts as an interface to Our World in Data datasets, allowing for an easy way to search through data used in over 3,000 charts and load them into the R environment.

The main function in owidR is `owid()`, which takes a chart id and returns a `tibble` (dataframe) of the corresponding OWID dataset. To search for chart ids you can use `owid_search()` to list all the chart ids that match a keyword or regular expression.

```{r human rights have changed over time}
owid_search("human rights")
```

Let’s use the human rights scores dataset.


```{r use the human rights dataset}
rights <- owid("human-rights-scores")
```

## Including Plots

`owid_plot()` makes it easy to visualise an owid dataset, plotting the first value column of an owid dataset. By default the mean score across all countries is plotted.

```{r visualize mean human rights score, echo=FALSE}
owid_plot(rights)
```

Use `summarise = FALSE` to show individual countries instead of the mean score. Unless a vector of entities is specified using the `filter` argument 9 random entities will be plotted. If the data is not a time-series then a bar chart of the entities values will be plotted.

```{r visualize human rights score across countries}
owid_plot(rights, summarise = FALSE, filter = c("North Korea", "South Korea", "France", "United Kingdom", "United States","Brazil","Spain"))

```

`owid_map()` makes it easy to create a choropleth world map of datasets that contain country level data. The Entities of the owid data must be country names. Currently the function plots data for the most recent year.

```{r Human right map, echo=FALSE}
owid_map(rights)
```
#### Creating Our World in Data style graphs

`owid_grapher()` creates graphs in the style of Our World in Data. The output of `owid_grapher()` can be piped into `grapher_line()` to add a line graph, into `grapher_map()` to add a world map, and into `grapher_labels()` to add labels to the graph. The graph is shown in the RStudio viewer, or when called in an RMarkdown html document is displayed within the document.

```{r}
rights %>% 
  owid_grapher(x = year, y = `Human Rights Score (Schnakenberg & Fariss, 2014; Fariss, 2019)`, 
               entity = entity) %>% 
  grapher_line(selected = c("North Korea", "South Korea", "France", "United Kingdom", "United States",
                            "Brazil","Spain")) %>% 
  grapher_map(palette = "RdYlGn", bins = c(-2, 0, 2, 4)) %>% 
  grapher_labels(title = "Human Rights Scores",
                 subtitle = "Values range from around -3.8 to around 5.4 (the higher the better)",
                 source = "Our World in Data; Schnakenberg and Fariss (2014); Fariss (2019)")
```

